---
title: "Effects of inflorescence architecture on within-individual variation in phenology and reproductive success of flowers in Lathyrus vernus"
subtitle: "Data preparation"
author : "Alicia Valdés"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    toc: yes
    toc_depth: 4
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
options(tibble.width = Inf)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r load packages, include=FALSE}
library(tidyverse)
library(readxl)
library(lubridate)
library(moments)
library(RColorBrewer)
library(ggthemes)
library(ggridges)
```

```{r Define ggplot themes and palettes, include=FALSE}
my_theme <- function(){
  theme_base()+theme(plot.background=element_rect(fill="white", colour=NA))+
  theme(legend.position="none")+theme(text=element_text(family="serif"))+
  theme(plot.title = element_text(hjust =-0.06))
}
my_theme_legend <- function(){
  theme_base()+theme(plot.background=element_rect(fill="white", colour=NA))+
  theme(text=element_text(family="serif"))+
  theme(plot.title = element_text(hjust =-0.06))
}
myPalette <- colorRampPalette(brewer.pal(11, "YlOrRd"))
```

# Read data for individual flowers from Excel files

```{r}
data_id_flowers_87 <- read_excel(
  "data/edited/individual_flower_characteristics_editsred.xlsx", 
                       sheet = "1987")
data_id_flowers_88 <- read_excel(
  "data/edited/individual_flower_characteristics_editsred.xlsx", 
                       sheet = "1988")
data_id_flowers_89 <- read_excel(
  "data/edited/individual_flower_characteristics_editsred.xlsx", 
                       sheet = "1989")
```

# Data prep individual flower data

```{r}
data_id_flowers_87 <- data_id_flowers_87 %>%
  rename(subplot=RUTA,number=GENET...2,
         shoot_id=SKOTT, # Empty when only one?
         raceme_id=KLASE,flower_id=BLOMMA,
         tot_racemes_in_shoot=TOTKLAS, # Total racemes of the shoot
         tot_flowers_in_raceme=TOTBLOM, # Total flowers of the raceme
         # Check that these two match with counts!
         fl_phen=BLOMFENO,fl_status=Blomstatus, # Seem equivalent
         n_ovules=OVULER,
         n_seeds=FR,n_pred_seeds=PREDERAD, # Check that this means this
         relpos_fl=Relpos,relpos_rac=Relraceme,
         opening_date=`New Phenoadj based on intervals`,
         fate_code=Framgångredef)%>%
  # 1 - aborted as flower
  # 2 - aborted as very small fruit
  # 3 - aborted as intermediate-sized fruit
  # 4 - aborted as large but not mature fruit
  # 5 - mature fruits with mature seeds
  # 6 - eaten/removed as flower or very small fruit
  # 7 - eaten/removed as intermediate or large fruit
  # All flowers that have seeds have the code 5
  dplyr::select(-Drecords,-FRAMGANG,-`Ny framgång`,-`Nyny framgång`,
                -AGG,-...19,-Success,-Phenolcorr,-Phenoladj,
                -GENET...27,-`Diff z-x`,-ID,
                -`New Phenomean for corrections based on intervals`)%>%
  # Remove empty columns and columns which I'm not sure would be useful
  # (maybe will use later)
  mutate(id = paste(subplot,number,sep=":"),year="1987",
         raceme_id=ifelse(is.na(raceme_id),4,raceme_id),
         # Fill raceme_id with 4 in two cases where it was missing
         # And total nr of racemes of the plant was 4
         relpos_fl=flower_id/(tot_flowers_in_raceme+1),
         relpos_rac=raceme_id/(tot_racemes_in_shoot+1)) # Recalculate
data_id_flowers_88 <- data_id_flowers_88 %>%
  rename(subplot=RUTA...1,number=GENET...2,
         shoot_id=SKOTT...3, # Empty when only one?
         raceme_id=KLASE...4,flower_id=BLOMMA...5,
         tot_racemes_in_shoot=TOTKLAS...6, # Total racemes of the shoot
         tot_flowers_in_raceme=TOTBLOM...7, # Total flowers of the raceme
         # Check that these two match with counts!
         fl_phen=BLOMFENO...8,fl_status=Blomstatus, # Seem equivalent
         n_ovules=OVULER,
         n_seeds=FR,n_pred_seeds=PREDERAD, # Check that this means this
         opening_date=`New Phenoadj based on intervals`,
         fate_code=Framgångredef...16)%>%
  dplyr::select(-`Phenocorrections based on intervals`,-Genet,
                -FRAMGANG...13,-`Ny FRAMGÅNG`,-`Nyny framgång`,
                -AGG,-Komm,-RUTA...22,-GENET...23,
                -SKOTT...24,-KLASE...25,-BLOMMA...26,-TOTKLAS...27,
                -TOTBLOM...28,-BLOMFENO...29,-FRAMGANG...30,-Check1,
                -Framgångredef...32,-Framgångsdiff,-...34) %>%
  # Remove empty columns and columns which I'm not sure would be useful
  # (maybe will use later)
  mutate(id = paste(subplot,number,sep=":"),year="1988",
         relpos_fl=flower_id/(tot_flowers_in_raceme+1),
         relpos_rac=raceme_id/(tot_racemes_in_shoot+1))%>%
  filter(!is.na(relpos_fl)) # Remove one row with all NAs
data_id_flowers_89 <- data_id_flowers_89 %>%
  rename(subplot=RUTA,number=GENET,
         shoot_id=SKOTT, # Empty when only one?
         raceme_id=KLASE,flower_id=BLOMMA,
         tot_racemes_in_shoot=TOTKLAS, # Total racemes of the shoot
         tot_flowers_in_raceme=TOTBLOM, # Total flowers of the raceme
         # Check that these two match with counts!
         fl_phen=BLOMFENO,fl_status=Blomstatus, # Seem equivalent
         n_ovules=OVULER,
         n_seeds=FRÖN,n_pred_seeds=PREDERAD, # Check that this means this
         opening_date=`Corrected pheno`,
         fate_code=Framgångredef)%>%
  dplyr::select(-Phenocorrection,-FRAMGANG,-`Ny framgång`,-`Nyny framgång`,
                -Genet,-HÅL,-Komm,-...22,
                -...23,-...24,-...25) %>%
  # Remove empty columns and columns which I'm not sure would be useful
  # (maybe will use later)
  mutate(id = paste(subplot,number,sep=":"),year="1989",
         relpos_fl=flower_id/(tot_flowers_in_raceme+1),
         relpos_rac=raceme_id/(tot_racemes_in_shoot+1))
```

```{r}
data_id_flowers<-full_join(
  full_join(data_id_flowers_87,data_id_flowers_88),
  data_id_flowers_89)%>%
  mutate(shoot_id=ifelse(is.na(shoot_id),1,shoot_id))%>%
  # Assign shoot_id=1 to all cases where NA
  mutate(fruit_set=factor(ifelse(fate_code==5,1,0)),
         initiated_fr=factor(ifelse(fate_code==2|fate_code==3|fate_code==4|
                                      fate_code==5|fate_code==7,1,0)),
         seed_set=n_seeds/n_ovules,
         prop_pred_seeds=round(n_pred_seeds)/n_seeds,
         seed_predation=ifelse(is.na(prop_pred_seeds),NA,
                               ifelse(n_pred_seeds>0,1,0)),
         year=factor(year))
```

```{r}
tot_shoots_in_plant<-data_id_flowers%>%group_by(year,id)%>%
  summarise(tot_shoots_in_plant=n_distinct(shoot_id))
tot_racemes_in_shoot<-data_id_flowers%>%group_by(year,id,shoot_id)%>%
  summarise(tot_racemes_in_shoot_1=n_distinct(raceme_id))
tot_flowers_in_raceme<-data_id_flowers%>%group_by(year,id,shoot_id,raceme_id)%>%
  summarise(tot_flowers_in_raceme_1=n_distinct(flower_id))
```

```{r}
data_id_flowers<-data_id_flowers%>%
  left_join(tot_shoots_in_plant)%>%
  left_join(tot_racemes_in_shoot)%>%
  left_join(tot_flowers_in_raceme)%>%
  dplyr::select(-tot_racemes_in_shoot,-tot_flowers_in_raceme)%>%
  rename(tot_racemes_in_shoot=tot_racemes_in_shoot_1,
         tot_flowers_in_raceme=tot_flowers_in_raceme_1)%>%
  # Substitute tot_racemes_in_shoot and tot_flowers_in_raceme
  # with my calculations, as there were some errors in the Excel file
  mutate(relpos_fl=flower_id/(tot_flowers_in_raceme+1),
         relpos_rac=raceme_id/(tot_racemes_in_shoot+1))
# Recalculate with the correct total values
```

# Transform dates

The dates are given in terms of four- or five-day intervals after the first recording. Convert them to calendar dates, then to julian dates, and then to number of days after the vernal equinox.

First create a table with information on each date for each year.

```{r}
dates <- tibble(year = as.factor(
  c(rep("1987",10),rep("1988",10),rep("1989",10))),
                date_num = rep(1:10,3), # Date in numeric format given in data
                date_calendar = as.Date(
                  ifelse(year==1987, seq(as.Date("1987-05-18"), 
                                         by = 4, length.out = 10),
                         # 1987: Start 18 May, 4-day intervals
                         ifelse(year==1988, seq(as.Date("1988-05-15"),
                                                by = 5, length.out = 10),
                                # 1988: Start 15 May, 5-day intervals
                                seq(as.Date("1989-05-07"), 
                                    by = 5, length.out = 10))),
                  # 1989: Start 7 May, 5-day intervals
                  # Calendar date
                  origin = "1970-01-01"),
  date_julian = yday(date_calendar), # Julian date
  date_vernal = ifelse(year==1987,date_calendar-as.Date("1987-03-21"),
                       ifelse(year==1988,date_calendar-as.Date("1988-03-20"),
                              date_calendar-as.Date("1989-03-20")))) 
# Days after vernal equinox
# Data on vernal equinox dates from https://data.giss.nasa.gov/ar5/srvernal.html
```

Calculate, for each year, the intercept and slope of the relationship among date_num (x)
and date_vernal or date_calendar (y).

```{r}
dates <- dates %>%
  group_by(year) %>%
  mutate(date_intercept_v = ifelse(year==1987,min(date_vernal)-4,
                                 min(date_vernal)-5),
         date_intercept_c = as.Date(ifelse(year==1987,min(date_calendar)-4,
                                 min(date_calendar)-5),origin = "1970-01-01"),
         date_slope_v = (date_vernal-date_intercept_v)/date_num,
         date_slope_c = (date_calendar-date_intercept_c)/date_num)
dates_summary <- summarise(dates, 
                           date_intercept_v = mean(date_intercept_v),
                           date_intercept_c = mean(date_intercept_c),
                           date_slope_v = mean(date_slope_v),
                           date_slope_c = mean(date_slope_c))
```

Add to individual flower data.

```{r}
data_id_flowers<-data_id_flowers%>%
  mutate(year=as.factor(year))%>%
  left_join(dates_summary, by = c("year" = "year"))%>%
  mutate(opening_date_c=date_slope_c*opening_date+date_intercept_c,
         opening_date_v=date_slope_v*opening_date+date_intercept_v,
         year=as.factor(year))%>%
  dplyr::select(-date_intercept_v,-date_intercept_c,-date_slope_v,-date_slope_c)
```

# Create unique ids for shoots, racemes and flowers

```{r}
data_id_flowers<-data_id_flowers%>%
  mutate(unique_shoot_id=paste(year,id,shoot_id,sep="_"),
         unique_raceme_id=paste(year,id,shoot_id,raceme_id,sep="_"),
         unique_flower_id=paste(year,id,shoot_id,raceme_id,flower_id,sep="_"))
```

# Save clean data as .csv

```{r}
write_csv(data_id_flowers,"data/clean/data_id_flowers.csv")
```

# Session info

```{r}
sessionInfo()
```

